Sequences complexity

### 2011-07-02 article.tex
Corrected a lot before writing last part

Using this complexity, one could define $1\ 2\ 2\ 3\ 3\ 3$ as 'count from one and repeat each number as much as itself', which is really easier to understand, memorize and extend for an human than a random sequence, like a telephone or a credit card number.
We will explore different ways to represent and manipulate such informations.

An other way to compute a list's complexity would be to use \textit{short term memory}, a list or tree containing last used elements.
An access to one of those elements would be its address for example binary coded.
This model is easily explainable considering that a human, if he was able to solve a problem using a specific operator, will naturally try it first when confronted to a new problem.
This list is implemented but not used because of its unpredictable character.

@@

\section{Tentatives}

@@

I defined a tuple \textit{operator} containing a field named \textit{times}.
Operator had the following methods, or \textit{generic functions} in Factor:
\begin{itemize}
  \item{cost, a function returning a operator's cost regarding its fields values. At this time, I wasn't thinking about short term memory.}
  \item{apply, a function which uses a digit and generate a sequence by uncompressing this operator on it.}
  \item{search, a function which tries to find current operator's application on a sequence.}
\end{itemize}

Given those methods, I defined three operators:
\begin{itemize}
  \item{A copy operator: using its predecessor's \textit{times}, copies given argument \textit{times} }
  \item{An increment operator: increments and copy by one given argument \textit{times}. If this argument is an operator, it will copy it by incrementing its \textit{times} field. }
  \item{A step operator: contains also a \textit{gap} argument and an \textit{operator} field, applies its operator by skipping as much elements as given by \textit{gap} on a list.}
\end{itemize}

The step operator was useful for this case:
$$1\  2\  2\  3\  3\  3\  4\  4\  4\  4$$
When compressed once using copy operator, it gave:
$$(copy\ times(1))\ 1\ (copy\ times(2))\ 2\ (copy\ times(3))\ 3\ (copy\ times(4))\ 4$$
Here, we can easily see that we have an increment on the digits and on the operators, but separated by one element each time. Using step operator, it gives:
$$(step\ operator\ times(4)\ gap(1)\ operator(increment\ times(1)))\ 1\ 2\ 3\ 4$$
The step operator will apply three times with a gap of one. As between each gap we only have one element, the increment operator in \textit{operator} field only applies once each time.
The step operator will apply on the directly following element, the copy operator.
A final compression would give:
$$(step\ operator\ times(4)\ gap(1)\ operator(increment\ times(1)))\ (increment\ times(4))\ 1$$

@@

This solution displeased Jean-Louis Dessalles, because of this strange step-operator not really 'natural', and its difficulties to understand and use.
That is why he asked me to only use two operators: copy and increment.
His point was that in the previous example, increment operator would have been applied at the same time on copy operator and on digits.
That is why I should put the argument into the operator instead of leaving it on the list.
Then, the increment operator would have seen a sequence of copy operator incrementing at the same time their \textit{times} and \textit{argument} fields.

@@

Now with this constraining consign, question was: "How can I apply different operators on the same object but on different fields?
I imagined a solution that I found pretty ugly, but functional: copy-operator would now both now its \textit{argument} and \textit{times}.
Increment-operator would moreover know where in his \textit{argument} it would apply: on its \textit{argument}, its \textit{times} or both.

Here is an application on previous example (first parenthesis contain \textit{argument} and second \textit{times}):
$$1\ 2\ 2\ 3\ 3\ 3\ 4\ 4\ 4\ 4$$
$$(increment(1)(1))\ (increment(2)(2))\ (increment(3)(3))$$
$$(increment(increment(1)(1))(3)(both))$$

@@

This solution is findable, putting apart the complexity problem (see below), and a huge problem which is at second iteration, how did he know that increment applied once applied only on times and not something else?
I choose 'times' because it has hence been able to compress a last time given that all increment operators applied on 'times', but I had no right to do so.

During a meeting with both Samuel Tardieu and Jean-Louis Dessalles, this solution worked pretty well, putting apart some complexity problems: sometimes 'good' -read 'expected'- solutions were not kept because they costed more than uncompressed sequence.
For example, with $1\ 2\ 2\ 3\ 3\ 4$, the program considered three increments of two times in array, but rejected it.
For the 'dirty' where-to-apply solution, Samuel cleverly suggested that instead of a restrictive three-choices field, I should use a indexes list which could also handle deepness indicating where the increment-operator should apply. This is the solution I adopted and customized in my final version.

@@

Current solution is axed on Factor's suffixed way: an operator is represented by a letter, $C$ for copy and $I$ for increment, and applies on a specified number of arguments \textit{before} it.
Elements in a compressed sequence are only compressed sequences, digits or operators letters.
Customized indexed work this way: increment operators own in their \textit{where} field a sequence of digits, which are indexes representing an element's absolute position in extended sequence.
For example, let us consider the following sequence:
$$\{\ 0\ 1\ \{\ 2\ \{\ 3\ 4\ \}\ 5\ \{\ 6\ \}\ \}\ 7\ \{\ 8\ \}\ 9\ \}$$
Digits represent each element's index.
Hence, any deepness level can be reached, and no need is to recall structural information with for example tree indexes, in which case element 4 would be accessed with 2.1.1 instead of just 4.
This notation is easier to manipulate and resulting \textit{where} sequence may even be compressed!
Jean-Louis Dessalles wasn't really satisfied by this index solution: it wouldn't respect Kantian structural form of the model (we should be able to manipulate shapes or vegetables as well and using them all along the model) and structure is absolutely not represented using it.
He suggested the use of masks, more 'natural' and compressed than my indexes notation, but it doesn't solve the structure's respect problem and is actually in bijection with my system (as it is more important to act simply and 'humanly' than saving RAM, we can neglect the compression difference).

For easier representation, an \textit{argument} or a \textit{where} field reduced to only one argument is not contained into a sequence, but inserted as is.

@@

As we saw before, when we find an increment operator used only once (\textit{times} = 1), its \textit{where} field doesn't mean anything, as when decompressed, \textit{argument} is only extracted, $I$, \textit{times} and \textit{where} being simply removed form the sequence.

@@

This is of course one over a lot of obtainable solutions, but it is findable by our process so we have the right to consider it.
Here we will make an 'error' which will cost us the solution:
$$\{\ 1\ 0\ 1\ I\ \}\ \textcolor{red}{0}\ 1\ I$$
$$\{\ 1\ 0\ 1\ I\ \}\ \textcolor{green}{2}\ 2\ I$$
$$\{\ 1\ 0\ 1\ I\ \}\ \textcolor{green}{2}\ 3\ I$$
$$\{\ 1\ 0\ 1\ I\ \}\ \textcolor{green}{2}\ 4\ I$$
Now, we would like to obtain
$$\{\ \{\ 1\ 0\ 1\ I\ \}\ \{\ 0\ 2\ \}\ 3\ I\ \}\ 6\ 4\ I$$
but our program will \textcolor{red}{fail} when comparing $0$ and $2$.

@@

An other way to improve the program is to solve the problem how to indicate where to apply the increment-operator in a natural, compressed and Kantian way.

This problem has been there since the beginning of the project, and even if a lots of plus or minus elegant solutions has been found, a lot of work is to been done on this subject.

\subsection{Short memory}

As explained before, the use of short memory is not reliable and quite difficult to implement.
Even though a whole implementation has been done, it is not used in final version because of its random characteristic.

@@

It revealed itself difficult to use at first but incredibly efficient for this particular subject.
The project itself proved to be harder than expected, because of its 'human' closely restriction.
It has been a opportunity to consider and try to bring solutions in a little known field.

>>>>

Using this complexity, one could define $1\ 2\ 2\ 3\ 3\ 3$ as 'count from one and repeat each number as much as itself', which is really easier to understand, memorize and extend for an human than a random sequence, like a telephone or credit card number.
We will explore different ways to represent and manipulate such informations.

An other way to compute a list's complexity would be to use \textit{short term memory}, a list or tree containing last used elements.
An access to one of those elements would be its binary coded address for example.
This model is easily explainable considering that a human, if he was able to solve a problem using a specific operator, will naturally try it first when confronted to a new problem.
This list is implemented but not used because of its unpredictable character.

@@

\section{Attempts}

@@

I defined a tuple \textit{operator} containing a field named \textit{times}.
Operator had the following methods, or \textit{generic functions} in Factor:
\begin{itemize}
  \item{cost, a function returning an operator's cost regarding its fields values.}
  \item{apply, a function which uses a sequence and generate a sequence by uncompressing this operator on it (often only using its first element).}
  \item{search, a function which tries to find current operator's application on a sequence.}
\end{itemize}

Given those methods, I defined three operators:
\begin{itemize}
  \item{A copy operator: using its predecessor's \textit{times} field, copies given argument \textit{times} times. }
  \item{An increment operator: increments and copy given argument \textit{times} times. If this argument is an operator, it will copy it by incrementing its \textit{times} field. }
  \item{A step operator: also contains a \textit{gap} argument and an \textit{operator} field, applies its operator by skipping as much elements as given by \textit{gap} on a list.}
\end{itemize}

The step operator was useful for this case:
$$1\  2\  2\  3\  3\  3\  4\  4\  4\  4$$
When compressed once using copy operator, it gave:
$$(copy\ times(1))\ 1\ (copy\ times(2))\ 2\ (copy\ times(3))\ 3\ (copy\ times(4))\ 4$$
Here, we can easily see that we have an increment on the digits and on the operators, but separated by one element each time. Using step operator, it gives:
$$(step\ operator\ times(4)\ gap(1)\ operator(increment\ times(1)))\ 1\ 2\ 3\ 4$$
The step operator will apply three times with a gap of one. As between each gap we only have one element, the increment operator in \textit{operator} field only applies once each time.
The step operator will apply on the directly following element, the copy operator.
A final compression would give:
$$(step\ operator\ times(4)\ gap(1)\ operator(increment\ times(1)))$$
$$(increment\ times(4))\ 1$$

@@

This solution displeased Jean-Louis Dessalles, because of this strange step-operator not really 'natural', and its difficulties to be understood and used.
That is why he asked me to only use two operators: copy and increment.
His point was that in the previous example, increment operator should have been applied at the same time on copy operator and on digits.
That is why I should put the argument into the operator instead of leaving it on the list.
Then, the increment operator would have seen a sequence of copy operator incrementing at the same time their \textit{times} and \textit{argument} fields.

@@

Now with this constraining consign, question was: "How can I apply different operators on the same object but on different fields?
I imagined a solution that I found pretty ugly, but functional: copy-operator would now both now its \textit{argument} and \textit{times}.
Increment-operator would inherit from copy-operator and moreover know where in his \textit{argument} it would apply: on its \textit{argument}, its \textit{times} or both. This information was stored on a \textit{where} field.

Here is an application on previous example (first parenthesis contain \textit{argument}, second \textit{times} and last \textit{where}):
$$1\ 2\ 2\ 3\ 3\ 3\ 4\ 4\ 4\ 4$$
$$(increment(1)(1)(arg))\ (increment(2)(2)(arg))\ (increment(3)(3)(arg))$$
$$(increment(increment(1)(1)(arg))(3)(both))$$

@@

This solution is findable, putting apart the complexity problem (see below), and a huge problem which is at second iteration, how did he know that increment applied once applied only on times and not something else?
I chose 'times' because it has hence been able to compress a last time given that all increment operators applied on 'times', but I had no right to do so.

During a meeting with both Samuel Tardieu and Jean-Louis Dessalles, this solution worked pretty well, putting apart some complexity problems: sometimes 'good' (think 'expected') solutions were not kept because they costed more than uncompressed sequence.
For example, with $1\ 2\ 2\ 3\ 3\ 4$, the program considered three increments of two times in array, but rejected it.
For the 'dirty' where-to-apply solution, Samuel cleverly suggested that instead of a restrictive three-choices field, I should use a indexes list which could also handle deepness indicating where the increment-operator should apply. This is the solution I adopted and customized in my final version.

@@

Current solution is axed on Factor's suffixed way: an operator is represented by a letter, $C$ for copy and $I$ for increment, and applies on a specified number of arguments \textit{before} it.
Elements in a compressed sequence are only compressed sequences, digits or operators letters.
Customized indexes work this way: increment operators own in their \textit{where} field a sequence of digits, which are indexes representing an element's absolute position in extended \textit{argument}.
For example, let us consider the following sequence:
$$\{\ a_0\ a_1\ \{\ a_2\ \{\ a_3\ a_4\ \}\ a_5\ \{\ a_6\ \}\ \}\ a_7\ \{\ a_8\ \}\ a_9\ \}$$
Indexes are used extended indexes.
Hence, any deepness level can be reached, and no need is to recall structural information, like for example with tree indexes, in which case element $a_4$ would be accessed with $2.1.1$ instead of $4$.
This notation is easier to manipulate and resulting \textit{where} sequence may even be compressed!

Jean-Louis Dessalles wasn't really satisfied by this index solution: it wouldn't respect Kantian structural form of the model (we should be able to manipulate shapes or vegetables as well and using them all along the model) and structure is absolutely not represented using it.
He suggested the use of masks, more 'natural' and compressed than my indexes notation, but it doesn't solve the structure's respect problem and is actually in bijection with my system (as it is more important to act simply and 'humanly' than saving RAM, we can neglect the compression difference).

For easier representation, an \textit{argument} or a \textit{where} field reduced to only one element is not contained into a sequence, but inserted as is.

We also define operators' arguments order:
\begin{itemize}
  \item{$WHAT\ TIMES\ C$}
  \item{$WHAT\ WHERE\ TIMES\ I$}
\end{itemize}

@@

As we saw before, when we find an increment operator used only once (\textit{times} = 1), its \textit{where} field doesn't mean anything, as when decompressed, \textit{argument} is only extracted, $I$, \textit{times} and \textit{where} fields being simply removed from the sequence.

@@

This is of course one over a lot of obtainable solutions, but it is findable by our process so we have the right to consider it.
Here we will make an 'error' which will cost us the solution:
$$\{\ 1\ 0\ 1\ I\ \}\ \textcolor{red}{0}\ 1\ I$$
$$\{\ 1\ 0\ 1\ I\ \}\ \textcolor{green}{2}\ 2\ I$$
$$\{\ 1\ 0\ 1\ I\ \}\ \textcolor{green}{2}\ 3\ I$$
$$\{\ 1\ 0\ 1\ I\ \}\ \textcolor{green}{2}\ 4\ I$$
Now, we would like to obtain
$$\{\ \{\ 1\ 0\ 1\ I\ \}\ 2\ 1\ I\ \}\ 6\ 4\ I$$
but our program will \textcolor{red}{fail} when comparing $0$ and $2$.

@@

An other way to improve the program is to solve the 'how to indicate where to apply the increment-operator in a natural, compressed and Kantian way' problem.

It has been there since the beginning of the project, and even if a lots of plus or minus elegant solutions have been found, a lot of work is to been done on this subject.

\subsection{Short term memory}

As explained before, the use of short memory is not reliable and quite difficult to implement.
Even though a whole implementation has been done, it is not used in final version because of its random behaviour.

@@

It revealed itself difficult to use at first but incredibly efficient for this particular subject.
The project itself proved to be harder than expected, because of its 'human' restriction.
It has been a opportunity to consider and try to bring solutions in a little known field.
